Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-21 cav2017/cav2017.tex
More refinements

The verification of a hybrid continuous/discrete time/space system over an unbounded time, requires the
non-deterministic simulation of an infinite set of states, which is undecidable. Our initial approach creates an
abstraction of this system by establishing bounds on the error between such a system and a discrete-time/space model.
This model is still however, quite complex, hard to verify, and many of the states involved are inconsequential to
the safety specification.
Our second approach, rather than look at a direct bi-simulation in discrete time, explores a much higher abstraction,
evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a
single iteration, in addition to a stability specification defined by Jury's criteria, and try to synthesise a controller that meets these specifications.
A second abstraction that encompasses the accelerated continuous reach-tube over an unbounded time is used to verify that the controller meets the full safety specification. If this step fails, we find a counterexample iteration that causes the selected system to fail and create a new constraint to refine our abstraction.
The advantage of this model is that it starts with a very simple description regardless of the complexity of the overall dynamics, and only expands to more complex models when a solution cannot be found.

@@

%
\begin{align}
\label{eq:observer_LTI_cf}
\vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right]
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\mat{T}\mat{B}_d&
\mat{B}_{tn}=&\mat{T}\mat{B}_d&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

\begin{align*}
\mat{A}_{t}=\mat{T}\mat{A}_d\mat{T}^{-1}&\left[
\begin{array}{ccccc}
0&1&0&\cdots&0\\
0&0&1&\cdots&0\\
\vdots&\vdots&\vdots&\ddots&\vdots\\
0&0&0&\cdots&1\\
k_n-a_n&k_{n-1}-a_{n-1}&k_{n-2}-a_{n-2}&\cdots&k_1-a_1
\end{array}\right]
\label{eq:cf_SISO_2}
\end{align*}

@@

For a stabilized closed-loop, we need these roots to remain within the
unit circle (left hand side of the plane for continuous time).  Whilst
this is not a sufficient condition for verifying safety properties, it
is a necessary one; therefore, we review this verification process.  
The direct approach would be to calculate the roots of the
polynomials (\emph{i.e.}, the matrix eigenvalues), but this ends
up being too costly, particularly in the case of synthesis, so we use
an indirect method.

Notice that in both our polynomials $c_0=1$ and we only use reals.
Now that we have established convergence, we may examine the safety 
specification of the state space. In the case of SISO systems, 
this is given by a range of the input and a range of the output, 
which are considered safe ($\vec{u}_k \in [\underline{u}\ \ \overline{u}] \wedge \vec{y}_k \in [\underline{y}\ \ \overline{y}]$). 
Whilst there are other properties, we may include in the verification, 
we will solely focus on these for simplicity of the explanation.
We will assume that we do not want saturation in our system, 
hence the control loop has only one mode. We have also identified 
the quantization noise as a bounded non-deterministic input.
First we must translate the safety specification into the state-space. 
Replacing on the second part of equation \eqref{eq:observer_LTI_cf} we obtain
%
\begin{equation}
\Psi := \underline{y}-\overline{\mat{B}\vec{u}}<\mat{C}_t\vec{x}_k<\overline{y}-\underline{\mat{B}\vec{u}}
\label{eq:of_bounds}
\end{equation}
Note that we use the bounds of the input after the transformation by $\mat{B}$ 
since the matrix may reverse the bounds. Eq.~\eqref{eq:of_bounds} can be solved 
using linear programming as it describes a polyhedral region in $\mathbb{R}^n$.
The second step is evaluating the progression of the system over time to calculate 
$\vec{x}_k$ to validate against this specification. For this step we require 
an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
\vec{x}=&\mat{A}_t^k\vec{x}_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}\left [\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] : \mat{E}\mat{T} \vec{x}_0<\mat{f}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{\vec{x} : \mat{E}\mat{T} \vec{x}_0<\mat{f} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \left[\begin{array}{c}\vec{\nu}_2\\ \vec{\nu}_1\end{array}\right] : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf}. It is worth noting that $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.

>>>>

The verification of a hybrid continuous/discrete system over an unbounded time, requires the
non-deterministic simulation of an infinite set of states, which is undecidable. Our initial approach creates an
abstraction of this system by establishing bounds on the error between such a system and a discrete-time/continuous-space model.
This model is still however, quite complex, hard to verify, and many of the states involved are inconsequential to
the safety specification.
Our second model, rather than look at a bisimulation in discrete time, explores a much higher abstraction,
evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specifcation.
We begin with a minimum set of linear inequalities describing the constraints caused by the input and the initial state during a
single iteration, in addition to a stability specification defined by Jury's criteria, and try to synthesise a controller that meets these specs.
A second abstraction that encompasses the accelerated continuous reach-tube over an unbounded time is used to verify that the controller meets the full safety specification. If this step fails, we find a counterexample iteration and initial state and create a new constraint to refine our abstraction.
The advantage of this model is that it starts with a very simple description regardless of the complexity of the overall dynamics, and only expands to more complex models when a solution cannot be found.

@@

%
\begin{align}
\label{eq:observer_LTI_cf}
\vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn} (\vec{\nu}_1 + \vec{\nu}_2 + \vec{\nu}_3)
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\mat{T}\mat{B}_d&
\mat{B}_{tn}=&\mat{T}\mat{B}_d&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

\begin{align*}
\mat{A}_{t}=\mat{T}\mat{A}_d\mat{T}^{-1}=&\left[
\begin{array}{ccccc}
0&1&0&\cdots&0\\
0&0&1&\cdots&0\\
\vdots&\vdots&\vdots&\ddots&\vdots\\
0&0&0&\cdots&1\\
k_n-a_n&k_{n-1}-a_{n-1}&k_{n-2}-a_{n-2}&\cdots&k_1-a_1
\end{array}\right]
\label{eq:cf_SISO_2}
\end{align*}

@@

For a stabilized closed-loop, which is necessary for our safety property, we need
these roots to remain within the unit circle, which we verify using Jury's criteria.
Notice that in our polynomials $c_0=1$ and we only use reals.

Once we have established convergence, we may examine the 
specification of the state and input spaces ($\phi_{safety} \wedge \phi_{input}$).
Whilst there are other properties, we may include in the verification, 
we will solely focus on these for simplicity of the explanation.
We assume that we do not want saturation in our system, 
hence the control loop has only one mode. We have also identified 
the quantization noise as a bounded non-deterministic input.
%
The first step constraints
\begin{equation}
\mat{K} \models \vec{u}_k=-\mat{K}\vec{x}_k : \phi_{safety} \wedge \phi_{input}
\label{eq:of_bounds}
\end{equation}
The second step is evaluating the progression of the system over time to evaluate 
$\vec{x}_{k+1} \models \phi_{safety}$. For this step we require 
an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
\vec{x}=&\mat{A}_t^k\vec{x}_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}(\vec{\nu}_1+\vec{\nu}_2+\vec{\nu}_3) : \vec{x}_0 \models \phi_{init}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{\vec{x} : \vec{x} \models \phi_{init} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \vec{\nu}_1+\vec{\nu}_2+ \vec{\nu}_3 : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \wedge \nu_3 \in \left[-q3\ \ q3\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf} and $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.
